Non-word repetition in Yélî Dnye

### 2020-09-17 manuscript.Rmd
add mic info; update to stimuli3.txt

- look through this manuscript for bits of text to be written by **mc**
- double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for... If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
- double check my frequency entries in segments.xlsx using the search function in https://phoible.org/parameters; take a look at http://phoible.github.io/conventions/ in case you see something about double articulation (I didn't find tp or lBj)
- double check also stim_seg_freq -- why are some items still with NA freq?
- draft discussion

@@

A second list of candidate items was generated in 2019 by selecting complex consonants and systematically crossing them with all the vowels in the Yélî inventory to produce consonant-vowel monosyllables. As before, items were automatically excluded if they appeared in the dictionary. Additionally, since hearing vowel length in monosyllables in isolation is challenging, any item that had a short/long real word neighbor was filtered out. Since the phonology and phonetics of Yélî is still in the process of being described [@levinsoninprepydgrammar], there could have been undocumented constraints that rendered items illegal. Therefore, we made sure that the precise consonant-vowel sequence occurred in some real word in the dictionary (i.e., that there was a longer word included the monosyllable as a subsequence). These candidates were presented to one informant, for a final check that they did not mean anything. Together with the 2018 selection, they were recorded using a headset **XXX **mc** please fill in** and an Olympus **XXX **mc** please fill in** from the written form presented together with the same item orally (by the second author). The complete recorded list was finally presented to two more informants, who could repeat all the items and who confirmed there were no real words. Even so, there was one monosyllable that was often identified as a real word (intended "yî" /yXX/; identified as "yi" /yi/, *tree*). This item is removed from analyses below.

@@

```{r tab2-stims, results = "asis"}
#add orthographic representations targets & other stims chars
stims.tbl <- read_delim("stimuli2.txt", delim = "\t") %>%
  arrange(factor(type, levels = c("practice", "1", "2", "3", "4"))) %>%
  select(type, target_ortho, phono) %>%
  rename("Type" = type, "Orthographic" = target_ortho, "Phonological" = phono)

@@

The child was donned a headset (**xx **mc** please fill in** for most of the children, SHURE WH20 XLR headset with a dynamic microphone for the rest), recorded into the left channel into a Tascam DR40x digital recorder. For most children, the headset could not stay comfortably on the child's head, and thus it was placed on the child's shoulders, with the microphone carefully placed close to the child's mouth. A local informant sat next to the child, to would provide the instructions and, if needed, coach the child to make sure, using the three practice items as well as real words, that they understood that the task was to repeat the items precisely without changing anything. An experimenter (the first author) delivered the elicitation stimuli to the local informant and the child over headphones.

@@

```{r Fig2-xling-freq, fig.pos = "!t", fig.align = "center", out.width = '65%', fig.cap="NWR scores for individual target words as a function of the average frequency with which each phone is found across languages."}
aggregate(trnsc$correct*100,by=list(trnsc$target,trnsc$avg_fr),mean)->xling
colnames(xling)<-c("Target","Avg_fr","Word")
stims.ipa <- read_delim("stimuli2.txt", delim = "\t") %>%
  select(target, target_ortho, phono)
xling <- merge(xling, stims.ipa, by.x = "Target", by.y = "target")

>>>>

- double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for... If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
- double check my frequency entries in segments.xlsx using the search function in https://phoible.org/parameters; take a look at http://phoible.github.io/conventions/ in case you see something about double articulation (I didn't find tp or lBj)
- double check also stim_seg_freq -- why are some items still with NA freq?
- draft discussion

@@

A second list of candidate items was generated in 2019 by selecting complex consonants and systematically crossing them with all the vowels in the Yélî inventory to produce consonant-vowel monosyllables. As before, items were automatically excluded if they appeared in the dictionary. Additionally, since hearing vowel length in monosyllables in isolation is challenging, any item that had a short/long real word neighbor was filtered out. Since the phonology and phonetics of Yélî is still in the process of being described [@levinsoninprepydgrammar], there could have been undocumented constraints that rendered items illegal. Therefore, we made sure that the precise consonant-vowel sequence occurred in some real word in the dictionary (i.e., that there was a longer word included the monosyllable as a subsequence). These candidates were presented to one informant, for a final check that they did not mean anything. Together with the 2018 selection, they were recorded using a Shure SM10A XLR dynamic headband microphone and an Olympus WS-832 stereo audio recorder (using an XLR to mini-jack adapter) from the written form presented together with the same item orally by the second author. The complete recorded list was finally presented to two more informants, who could repeat all the items and who confirmed there were no real words. Even so, there was one monosyllable that was often identified as a real word (intended "yî" /yXX/; identified as "yi" /yi/, *tree*). This item is removed from analyses below.

@@

```{r tab2-stims, results = "asis"}
#add orthographic representations targets & other stims chars
stims.tbl <- read_delim("stimuli3.txt", delim = "\t") %>%
  arrange(factor(type, levels = c("practice", "1", "2", "3", "4"))) %>%
  select(type, target_ortho, phono) %>%
  rename("Type" = type, "Orthographic" = target_ortho, "Phonological" = phono)

@@

The child was donned a headset (a Shure SM10A or WH20 XLR with a dynamic microphone on a headband, most children using the former), recorded into the left channel into a Tascam DR40x digital recorder. For most children, the headset could not stay comfortably on the child's head, and thus it was placed on the child's shoulders, with the microphone carefully placed close to the child's mouth. A local informant sat next to the child, to would provide the instructions and, if needed, coach the child to make sure, using the three practice items as well as real words, that they understood that the task was to repeat the items precisely without changing anything. An experimenter (the first author) delivered the elicitation stimuli to the local informant and the child over headphones.

@@

```{r Fig2-xling-freq, fig.pos = "!t", fig.align = "center", out.width = '65%', fig.cap="NWR scores for individual target words as a function of the average frequency with which each phone is found across languages."}
aggregate(trnsc$correct*100,by=list(trnsc$target,trnsc$avg_fr),mean)->xling
colnames(xling)<-c("Target","Avg_fr","Word")
stims.ipa <- read_delim("stimuli3.txt", delim = "\t") %>%
  select(target, target_ortho, phono)
xling <- merge(xling, stims.ipa, by.x = "Target", by.y = "target")
